Redwine quality Analysis by Gangadhara Naga Sai
========================================================

```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.
# install.packages('MASS')
# install.packages('memisc')
# install.packages('COUNT')
 #install.packages('corrplot')
# install.packages('sandwich')
# install.packages('msme')
# install.packages('GGally')
# install.packages('dplyr')
#install.packages('RGraphics')
library(RGraphics)
library(corrplot)
library(gridExtra)
library(RColorBrewer)
library(lattice)
library(MASS)
library(memisc)
library(COUNT)
library(msme)
library(sandwich)
library(GGally)
library(dplyr)
library(GlobalOptions)
#install.packages('GlobalOptions')
library(graphics)
library(ggplot2)
# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```

```{r , Load_the_Data  and data summary}
# Load the Data
# download.file("https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityReds.csv","wineQualityReds.csv",method="curl")
redwines<-read.csv('wineQualityReds.csv')
redwines$X<-NULL
print("Names of variables ")
names(redwines)
print("Dimensions of wine data")
dim(redwines)
print("Structure of wine data")
str(redwines)
print("Summary of Redwine data")
summary(redwines)
```

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}


```

```{r,echo=FALSE,message=FALSE,warning=FALSE}
transparent_theme <- theme(
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 axis.text.x = element_blank(),
 axis.text.y = element_blank(),
 axis.ticks = element_blank(),
 panel.grid = element_blank(),
 axis.line = element_blank(),
 panel.background = element_rect(fill = "transparent",colour = NA),
 plot.background = element_rect(fill = "transparent",colour = NA))

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
HistoBox <- function(x,i,l, na.rm = TRUE, ...) {
  m<-names(x)
  heading<-paste("Histogram and Boxplot of ",m[i])
  
  #Histogram of the variable x
  p1 <- ggplot(x,aes(x[,i],y=..density..))+
    geom_histogram(colour="black",alpha=1,fill="#993366",bins=30)+
    geom_line(stat='density',color="black",linetype = 2)+ ggtitle(heading) + 
     theme(plot.title = element_text(lineheight=.8, face="bold"))+
    theme( axis.title.x = element_blank())
  
  
  # Box plot of the x variable
  
  
p2 <- ggplot(redwines, aes(factor(1), x[,i]))+
  geom_jitter(alpha = 0.2,color='#993366')+
  geom_boxplot(alpha=0.1,color='black')+
  stat_summary(fun.y=mean,shape=1,col='red',fill='red',geom='point')+
  geom_hline(yintercept = (median(x[,i]) + sd(x[,i])*3.5),linetype = 4)+
  coord_flip()+
  labs(y =m[i])
  #+transparent_theme
  
  #text or comments
  
  text<-paste(l[i],"Where the Standard deviation is",round(sd(x[,i]),3),",beyond ",round(sd(x[,i]) * 3.5+median(x[,i]),3)," we will find outliers ","and the average from boxplot is ",round(mean(x[,i]),3),". Histogram highest value"," at ",m[i]," equal to ",Mode(x[,i]),".")
    
  p3 <- splitTextGrob(text)
    
  return(grid.arrange(p1,p2,p3))
}



```



```{r, }



# plots = lapply(seq_along(m), function(x) HistoBox(redwines[,x],comments))
# require(gridExtra)
# do.call(grid.arrange,  c(plots,ncol=1))

#lapply(seq_along(nm), function(i) HistoBox(redwines,i,comments))


```

```{r,message=FALSE,echo=FALSE,results='hide'}
comments <-list("Fixed acidity is slightly a skewed Normal distribution."
                ," Volatile acidity is similar to fixed acidity, also has a long tail distribution.It is having a bimodal distribution."
                ,"Citric acid has quite a long-tail distribution.But it is not a Normal Distribution.And there are no wine of 8% concentration as in data"
                ,"Outliers are many fot Residual sugar which  has a very long-tail distribution.Maximum values are in between 1.9 to 2.6."
                ,"Almost similar to residual sugar with a very long-tail distribution.Maximum of values are between 0.07 and 0.09"
                ,"Free sulfur dioxide has a positively skewed distribution. Maximum of the values are between 9.50 and 11.10."
                ,"Same as last one, lesser outliers and its interquartile range  seems not quite large. It also has a long-tail distribution. "
                ,"Apperar quite like a normal  distribution,most of the values falling between 0.995 and 1."
                ,"pH is also normal distribuition with most of the pH values falling between 3.1 and 3.5. Not much outliers are seen."
                ," For sulphites most of the values seems to be concentrated at 0.6. It is a long tailed distribution and quite similar to chlorides and residual sugar."
                ," Most wines have less than 11% alcohol. Alcohol has a long-tail distribution, with there only being a few outliers"
                ,"Similar to fixed acidity's distibution,since fixed acidity comprises the most.Total acidity values typically fall between 7 and 10 and there are very few outliers."
                ,"combined sulfur dioxide has a distribution similar to total sulfur dioxide."
                )
redwinesx<-redwines
redwinesx$total.acidity<-redwines$fixed.acidity+redwines$volatile.acidity
redwinesx$combined.sulphur.dioxide<-redwines$total.sulfur.dioxide-redwines$free.sulfur.dioxide
redwinesx<-redwinesx[,c(1:11,13,14,12)]
lapply(1:13, function(i) HistoBox(redwinesx,i,comments))
ggplot(aes(quality), data = redwines) +geom_bar(colour="black",alpha = 1,fill =  "#993366")

```

Quality ranges from 0 to 10,but in data minimum is 3 and maximum is 8, which means that most of the wines we will look at in the analysis are average wines.
```{r}
summary(redwines$quality)

```




# Univariate Analysis

### What is the structure of your dataset?
>Data set is regarding the wine quality and several chemical componets that it contains.there ae 1599 samples of wine with 10 variables(fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfer dioxide, density, pH, sulphates, alcohol, quality) of type numeric and 1 rating factor quality of type int.  

### What is/are the main feature(s) of interest in your dataset?
>Quality is the main feature of insterest ,given by 3 wine experts according to their knowledge and experience.Quality ranges from 0 to 10 but our data has least quality of 3 and highest quality of 8. Lets find out what are the main deciding factors for high quality wine.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
>There can lot more features since in real world so many factors affect the quality of Red wine.
>>Type of grapes used
>>flavor (like combination of different ingredients)
>>Color
>>taste(sweet,sour,bitter,etc)
>>total cost from the ingredients to final production of wine(since cost matters since high quality wine with less cost really matters)

### Did you create any new variables from existing variables in the dataset?
>Yes , i made total.acidity and combined.sulphur.dioxide, which may show some unseen trends.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
>Volatile acidity is having a bimodal distribution and Citric acid has quite a long-tail distribution.But it is not a Normal Distribution.
>the data was already tidy so there was no requirement for any adjustment.



# Bivariate Plots Section
## By using correlation we can find out some important insights among variables

```{r echo=FALSE, Bivariate_Plots}

#corrplot.mixed(cor(redwines))
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(cor(redwines), method="color", col=col(200),  
          order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
          sig.level = 0.01, insig = "blank", type="upper",
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
c<-cor(redwines)
print("Correlation among the variables")
sort(c[,12])


```

>Observing the  correlation, alcohol and volatile acidity, have a higher correlation with the quality of wine.Suphates and citric acid are also correlated with the quality of wine.
>Residual sugar has almost no correlation with quality.


```{r message=FALSE, echo=FALSE, results='hide'}
Boxinscatter <- function(x,i,l,factor, na.rm = TRUE, ...) {
  m<-names(x)
 heading<-paste(" BoxinScatter: \n",m[i]," vs ",factor)
  
  #Histogram of the variable x
  p1 <-ggplot(aes(factor(x[,factor]), x[,i]), data = x) +
    geom_jitter( alpha = .5,color="#993366")  +
    geom_boxplot( alpha = .2,color = 'black')+
    stat_summary(fun.y = "median",geom = "point",color = "red",shape = 18,size = 4)+
    geom_smooth(aes(x[,factor]-2, x[,i]),method = "lm",se = FALSE,size=1,color = 'red',linetype=2) +
    geom_hline(yintercept = (median(x[,i]) + sd(x[,i])*3.5),linetype = 4) +
    ylab(m[i])+xlab(factor)+theme(text = element_text(size=13))+ ggtitle(heading) + 
     theme(plot.title = element_text(lineheight=.8, face="bold",size = 15))
  
  a<-paste(with(data=x,cor.test(x[,i],x[,factor])))
  b<-paste(l[i],a[7]," with ","t value equal to",round(as.numeric(a[1]),3),", 95 percent confidence interval is",a[9],"with a correlation of",round(as.numeric(a[4]),3))
  
  
  p3 <- splitTextGrob(b,hjust = "right",vjust = "bottom",gp=gpar(fontsize=12),x = unit(0, "npc"))
  
  return(grid.arrange(p1,p3,heights=c(3,1)))
  
}
#a<-paste(with(data=x,cor.test(x[,i],x[,factor])))
commentz<-list("There positive correlation between fixed acidity and quality. The outliers fall in wine of quality 7 and 5."
               ,"We can see Negitive correlation is the trend here."
               ,"It is clear that  more citric acid higher the quality ,There are no wines of quality 8 that have no citric acid."
               ," Their is a very weak correlation between residual sugar and quality of wine shows here. Its correlation is also very close to zero."
               ,"A weak negetive correlation is seen."
               ,"There is not quite a trend seen as the median increases and decreases.It shows free.sulphur.dioxide is not good indicator of quality."
               ,"Oberving a negetive correlation.The outliers are present at quality 7,  less total sulfur dioxide leads to better wine."
               ,"A weak negitive correlation is seen.Trend between quality of wine and density goes down."
               ,"Again negitive correlation as pH is low it is more acidic, so the wine with lower pH seems to have good quality"
               ,"Positive correlation, but for best quality suphate should be more than 0.5 and less than 1.5."
               ,"A very high correlation is seen ,higher alcohol content will give high quality  wine."
               ,"A weak positive correlation is seen and very similar to fixed acidity."
               ,"It seems there is a negitive correlation ,but the median increases and decreases.")

#Boxinscatter(redwinesx,13,commentz,'quality')
lapply(1:13, function(i) Boxinscatter(redwinesx,i,commentz,'quality'))

   # Boxinscatter(redwines,9)
```

```{r}

scatter_smooth<- function(df,i,l,factor, na.rm = TRUE, ...){
  m<-names(df)
  heading<-paste(" Scatter_smooth: \n",factor," vs ",m[i])
 a<-paste(with(data=df,cor.test(df[,i],df[,factor])))
b<-paste(l[i],a[7]," with ","t value equal to",round(as.numeric(a[1]),3),", 95 percent confidence interval is",a[9],"with a correlation of",round(as.numeric(a[4]),3))

grid.arrange(ggplot(aes(x=df[,i],y=df[,factor]),data=df) +
  geom_point(alpha=0.3,color="#993366") +
    geom_smooth()+ylab(factor)+xlab(m[i])+
    theme(text = element_text(size=13))+ ggtitle(heading) + 
     theme(plot.title = element_text(lineheight=.8, face="bold",size = 15))
  ,splitTextGrob(b,hjust = "right",vjust = "bottom",gp=gpar(fontsize=12),x = unit(0, "npc")),heights=c(3,1))

}




```

```{r message=FALSE, echo=FALSE, results='hide'}
commentz<-list("From the graph chlorides ranges from 0.07 to 0.09, the density is positively correlated with chlorides."
               ,"The density of wine is inversely proportional to the alcohol present in it."
               ,"From the graph citric.acid ranges from 1.9 to 2.6, the density is positively correlated with citric acid."
               ,"The density is positively correlated with residual sugar."
)
redensity<-redwines[,c("chlorides","alcohol","citric.acid","residual.sugar","density")]
lapply(1:4, function(i) scatter_smooth(redensity,i,commentz,'density') )

```

```{r message=FALSE, echo=FALSE, results='hide'}
commentz<-list("pH is positively correlated with  volatile.acidity."
               
               ,"From scatter plot  high negative correlation coefficient is seen, pH of wine is inversely proportional to the fixed acidity."
               ,"pH is negatively correlated with total acidity of wine.")
redpH<-redwinesx[,c("volatile.acidity","fixed.acidity","total.acidity","pH")]
lapply(1:3, function(i) scatter_smooth(redpH,i,commentz,'pH') )
```

```{r message=FALSE, echo=FALSE, results='hide'}
commentz<-list("citric.acid is inversely related with volatile.acidity."
               ,"citric.acid is positively correlated with fixed.acidity. More the citric acid is added to the wine,fixed acidity increases."
               ,"Very similar to fixed acidty.")
redcitric<-redwinesx[,c("volatile.acidity","fixed.acidity","total.acidity","citric.acid")]
lapply(1:3, function(i) scatter_smooth(redcitric,i,commentz,'citric.acid') )

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
> Quality being the feature of interest,the correlation  between quality and each different variable in the dataset are examined.Quality of wine is directly proportional to the alcohol content and volatile acidity and inversely proportional to density,total sulfur dioxide and chlorides.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
> pH and volatile acidity are positively correleated, Higher the pH value means less acidity, but from plots a higher volatile acidity means more acidity.
>Density of wine has high negative correlation with the amount of alcohol  in wine.
>I was expecting a close relation between sulphur and sulphur dioxide,there seems no relation
with  correlation coefficient of 0.04. 

### What was the strongest relationship you found?
> correlation of quality with other variables
> volatile.acidity      total.sulfur.dioxide           density           chlorides 
         -0.39055778          -0.18510029          -0.17491923          -0.12890656 
                  pH  free.sulfur.dioxide       residual.sugar        fixed.acidity 
         -0.05773139          -0.05065606           0.01373164           0.12405165 
         citric.acid            sulphates              alcohol              quality 
          0.22637251           0.25139708           0.47616632           1.00000000 
> From the correlations we can clearly see alcohol positiely and volatile.acidity negitively  are having a strong relation with quality.
>And density and fixed acidity have a strong correlation.



# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
scatter<- function(df,factor1, factor2){
  #  m<-names(df)
  #   heading<-paste(" Scatter_smooth: \n",factor," vs ",m[i])
  #  a<-paste(with(data=df,cor.test(df[,i],df[,factor])))
  # b<-paste(l[i],a[7]," with ","t value equal to",round(as.numeric(a[1]),3),", 95 percent confidence interval is",a[9],"with a correlation of",round(as.numeric(a[4]),3)) 
  p1<-ggplot(aes(df[,factor1], df[,factor2], color = factor(quality)), data = df) +
    geom_jitter(size = 1.5,alpha = .5,fill='#993366')+
    scale_color_brewer(type="seq",guide=guide_legend(title='Quality'),palette=11)+
     geom_smooth(se = FALSE)+xlab(factor1) +
    ylab(factor2) +  theme(panel.background = element_rect( fill = "grey80"))
  
  print(ggplot(aes(df[,factor1], df[,factor2], color = factor(quality)), data = df) +
    geom_jitter(size = 1.5) +
  scale_color_brewer(type="qual"))
  
  p1+facet_wrap(~quality,ncol=2)  
  #grid.arrange(p1)
}
#?scale_color_brewer
# ?ggplot
# ?geom_smooth
scatter(redwines,"volatile.acidity","alcohol")
scatter(redwines$alcohol, "alcohol", redwines$volatile.acidity, "volatile acidity",11)
lapply(1:8, function(i) scatter(redwines,"volatile.acidity","alcohol",i))


```


```{r}
scatter<- function(df,factor1,i){
 
 drops <- c(factor1)
df1<-df[ , !(names(df) %in%  drops)]

 
 m<-names(df1)
#   heading<-paste(" Scatter_smooth: \n",factor," vs ",m[i])
#  a<-paste(with(data=df,cor.test(df[,i],df[,factor])))
# b<-paste(l[i],a[7]," with ","t value equal to",round(as.numeric(a[1]),3),", 95 percent confidence interval is",a[9],"with a correlation of",round(as.numeric(a[4]),3)) 
p1<-ggplot(aes(df[,factor1], df1[,i], color = factor(quality)),data=df1) +
  geom_jitter(size = 1.5, alpha = .3,color='#993366') +
  scale_color_brewer(type='seq', 
                     guide=guide_legend(title='Quality'),
                     palette = 11) +
  geom_smooth(se = FALSE) +
    xlab(factor1) +
    ylab(m[i]) +  theme(panel.background = element_rect( fill = "grey80"))


  p1+facet_wrap(~quality,ncol=2)
 # grid.arrange(p1)
}

#?theme
# ?ggplot
# ?geom_smooth
##scatter(redwines$alcohol, "alcohol", redwines$volatile.acidity, "volatile acidity",11)
scatter(redwines, "alcohol",10)
lapply(1:10, function(i) scatter(redwines, "alcohol",i))
redwines1<- redwines[,!c("alcohol")]
drops <- c("alcohol","quality")
redwines1<-redwines[ , !(names(redwines) %in%  drops)]
redwines1
redwines<-redwines[ ,  drops]
redwines1
refcols <- c("Trader", "System")
 #
 trading <- trading[, c(refcols, setdiff(names(trading), refcols))]
 names(trading)
dim(redwines1)
ggplot(aes(redwines$fixed.acidity, redwines$density, color = factor(quality)),data=redwines1) +
  geom_jitter(size = 1.5, alpha = .3,color='#993366')+facet_wrap(~quality,ncol=2) + geom_smooth(se = FALSE)
redwines1

```


```{r}
scatter<- function(df,factor1,i){
  #  m<-names(df)
  #   heading<-paste(" Scatter_smooth: \n",factor," vs ",m[i])
  #  a<-paste(with(data=df,cor.test(df[,i],df[,factor])))
  # b<-paste(l[i],a[7]," with ","t value equal to",round(as.numeric(a[1]),3),", 95 percent confidence interval is",a[9],"with a correlation of",round(as.numeric(a[4]),3)) 
  p1<-ggplot(aes(df[,factor1], df[,i], color = factor(quality)), data = df) +
    geom_jitter(size = 1.5,alpha = .5,fill='#993366')+
    scale_color_brewer(type="seq",guide=guide_legend(title='Quality'),palette=11)+
     geom_smooth(se = FALSE)+xlab(factor1) +
    ylab(df[,i]) +  theme(panel.background = element_rect( fill = "grey80"))
  
  p2<-ggplot(aes(df[,factor1], df[,i], color = factor(quality)), data = df) +
    geom_jitter(size = 1.5) +
  scale_color_brewer(type="qual")
  
  p0<-p1+facet_wrap(~quality,ncol=2)  
  print(p0)
  print(p2)
  
  #grid.arrange(p1)
}
#?scale_color_brewer
# ?ggplot
# ?geom_smooth
#scatter(redwines,"volatile.acidity","alcohol")
#scatter(redwines$alcohol, "alcohol", redwines$volatile.acidity, "volatile acidity",11)
lapply(1:8, function(i) scatter(redwines,"volatile.acidity",i))

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
